Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1)

### 2014-04-25 WSSSPE1_report/WSSSPE1_summary_report.tex
Section 8: Fiddling.
changes to 5

Table~\ref{tab:defining-sustainability} reproduces a summary created
by C.~Lenhardt ~\cite{lenhardt-wssspe1-panel}, which shows for each
contribution to the Modeling Sustainability panel what it meant when
referring to software, how it defined software sustainability, and
which approach it suggested to understand or evaluate sustainability.
\begin{table}[t]
  \begin{scriptsize}
    \begin{center}
      \caption{Summary of Modeling Sustainability papers from Policy Panel.  Adapted from~\cite{lenhardt-wssspe1-panel}.}
      \label{tab:defining-sustainability}
      \begin{tabular}{|p{2.3cm}|p{3.6cm}|p{4.4cm}|p{4.8cm}|}
                \hline
{\bf Paper/Authors}
& {\bf Software}
& {\bf Sustainability}
& {\bf Approach to Understand or Evaluate Sustainability} \\
                \hline
Calero, et al.~\cite{Calero_WSSSPE}
& General notion of software. Not explicitly defined.
& Sustainability is linked to quality.
& Add to ISO \\
                \hline
Venters, et al.~\cite{Venters_WSSSPE}
& Software as science software; increasingly complex; service-oriented computing
& Extensibility, interoperability, maintainability, portability, reusability, scalability, efficiency
& Use various architecture evaluation approaches to assess sustainability \\
                \hline
Pierce, et al.~\cite{Pierce_WSSSPE}
& Cyberinfrastructure software
& Sustainable to the extent to which there is a community to support it
& Open community governance \\
                \hline
Katz, et al.~\cite{Katz_WSSSPE}
& E-research infrastructures (i.e. cyberinfrastructure)
& Persisting over time, meeting original needs and projected needs
& Equates models for the creation of software with sustaining software \\
                \hline
Lenhardt,~et~al.~\cite{Lenhardt_WSSSPE}
& Broadly defined as software supporting science
& Re-use; reproducible science
& Comparing data management life cycle to software development life cycle \\
                \hline
Weber, et al.~\cite{Weber_WSSSPE}
& Software broadly defined; a software ecosystem
& Software niches
& Ecological analysis and ecosystem \\
                \hline
     \end{tabular}
    \end{center}
  \end{scriptsize}
\end{table}

@@

A second theme in these papers was the continued availability of
resources to accomplish the goals of sustainability. The elements of
these frameworks focused far more on the organization of a software
project than they did on characteristics of the artifact itself
(although it is certainly true that the adjectives discussed above
could be applied to a software project). For example Pierce et
al.~\cite{Pierce_WSSSPE} focus on the way that the project is run,
particularly in terms of how those involved communicate and jointly
set priorities, a process they call governance. In particular they
argue that because sustainability is related to having ongoing resources,
governance must be open to receive diverse input (by occurring online,
asynchronously, or publicly) and thus have the potential to
``transform passive users into active contributors.'' They argue that
the Apache Software Foundation's incubation process teaches this and
could be learned from by projects throughout scientific software.
Katz and Proctor~\cite{Katz_WSSSPE} also discuss governance,
describing two modes: ``top-down'' and ``bottom-up'' governance.  They
place governance alongside questions of the artifact (``technical''),
and questions of the origins of resources, in terms of who is funding
the work surrounding a piece of software (``the political'') and the
manner in which resources come to the project initially (commercial,
open source, closed partnerships, grant funded) and long-term (all
four plus paid support).

@@

One such example is the \toolname{Galaxy} project~\cite{Galaxy}. It
follows agile software development practices and implements standards
such as test-driven development and socialized bug managing practices
via \toolname{trello}. Galaxy \emph{histories} and \emph{toolshed}
offers easy community sharing of data and tools further promoting a
collaborative environment. The project closely follows the
guidelines described in Carver and Thiruvathukal~\cite{Carver_WSSSPE}
and many from Prli\'{c} and Procter~\cite{Prlic_WSSSPE}. Many utility
software projects are often developed aiming better utilizing a
particular, new infrastructure and architecture, e.g.,
\toolname{MVAPICH}, \toolname{VisIT}, \toolname{ParaView}. Similarly, to leverage the power
of emerging architectures such as accelerators, new code and libraries
are required. The experience of one such effort as described in
Ferenbaugh~\cite{Ferenbaugh_WSSSPE} which met with a limited success
but nonetheless with many invaluable lessons were learned about
influential cultural and technical aspects in sustainable software
development practices.

>>>>

Table~\ref{tab:defining-sustainability} is based on a summary 
by Lenhardt ~\cite{lenhardt-wssspe1-panel}, which shows for each
contribution to the Modeling Sustainability panel what it meant when
referring to software, how it defined software sustainability, and
which approach it suggested to understand or evaluate sustainability.
\begin{table}[t]
  \begin{scriptsize}
    \begin{center}
      \caption{Summary of Modeling Sustainability papers from Policy Panel.  Adapted from~\cite{lenhardt-wssspe1-panel}.}
      \label{tab:defining-sustainability}
      \begin{tabular}{|p{2.3cm}|p{3.6cm}|p{4.4cm}|p{4.8cm}|}
                \hline
{\bf Paper/Authors}
& {\bf Software}
& {\bf Sustainability}
& {\bf Approach to Understand or Evaluate Sustainability} \\
                \hline
Calero, et al.~\cite{Calero_WSSSPE}
& General notion of software. Not explicitly defined.
& Sustainability is linked to quality.
& Add to ISO \\
                \hline
Venters, et al.~\cite{Venters_WSSSPE}
& Software as science software; increasingly complex; service-oriented computing
& Extensibility, interoperability, maintainability, portability, reusability, scalability, efficiency
& Use various architecture evaluation approaches to assess sustainability \\
                \hline
Pierce, et al.~\cite{Pierce_WSSSPE}
& Cyberinfrastructure software
& Sustainable to the extent to which there is a community to support it
& Open community governance \\
                \hline
Katz, et al.~\cite{Katz_WSSSPE}
& E-research infrastructures (i.e. cyberinfrastructure)
& Persisting over time, meeting original needs and projected needs
& Equates models for the creation of software with sustaining software \\
                \hline
Lenhardt,~et~al.~\cite{Lenhardt_WSSSPE}
& Broadly defined as software supporting science
& Re-use; reproducible science
& Comparing data management life cycle to software development life cycle \\
                \hline
Weber, et al.~\cite{Weber_WSSSPE}
& Software broadly defined; a software ecosystem
& Software niches
& Ecological analysis and ecosystem \\
                \hline
     \end{tabular}
    \end{center}
  \end{scriptsize}
\end{table}

@@

A second theme in these papers was the continued availability of
resources to accomplish the goals of sustainability. The elements of
these frameworks focused far more on the organization of a software
project than they did on characteristics of the artifact itself
(although it is certainly true that the adjectives discussed above
could be applied to a software project). For example Pierce et
al.~\cite{Pierce_WSSSPE} focus on the way that the project is run,
particularly in terms of how those involved communicate and jointly
set priorities, a process they call governance. In particular they
argue that because sustainability is related to having ongoing resources,
governance must be open to receive diverse input (by occurring online,
asynchronously, or publicly) and thus have the potential to
``transform passive users into active contributors.'' They argue that
the Apache Software Foundation's incubation process teaches this and
could be learned from by projects throughout scientific software.
Katz and Proctor~\cite{Katz_WSSSPE} also discuss governance,
describing two modes: ``top-down'' and ``bottom-up'' governance.  They
place governance alongside technical questions about the software,
political questions about who is funding
the work surrounding the software, and the
manner in which resources come to the project both initially (commercial,
open source, closed partnerships, grant funded) and long-term (all
four plus paid support).

@@

One such example is the \toolname{Galaxy} project~\cite{Galaxy}. It
follows agile software development practices and implements standards
such as test-driven development and socialized bug managing practices
via \toolname{trello}. \toolname{Galaxy} \emph{histories} and \emph{toolshed}
offer easy community sharing of data and tools further promoting a
collaborative environment. The project closely follows the
guidelines described in Carver and Thiruvathukal~\cite{Carver_WSSSPE}
and many from Prli\'{c} and Procter~\cite{Prlic_WSSSPE}. Many utility
software projects are often developed aiming better utilizing a
particular, new infrastructure and architecture, e.g.,
\toolname{MVAPICH}, \toolname{VisIT}, \toolname{ParaView}. Similarly, to leverage the power
of emerging architectures such as accelerators, new code and libraries
are required. The experience of one such effort as described in
Ferenbaugh~\cite{Ferenbaugh_WSSSPE} which met with a limited success
but nonetheless with many invaluable lessons were learned about
influential cultural and technical aspects in sustainable software
development practices.
